Modelling BCG vaccination in the UK

### 2019-08-06 rmds/02-background.Rmd
updated getTBinR figures

```{r plot-incidence, echo = FALSE, warning = FALSE, message = FALSE, dpi = 300, fig.height = 12, fig.width = 12, fig.scap = "TB notifications in England and Wales from 1913 to 2017, stratified initially by respiratory/non-respiratory status and from 1982 by pulmonary/non-pulmonary TB.", fig.cap = "TB notifications in England and Wales from 1913 to 2017, stratified initially by respiratory/non-respiratory status and from 1982 by pulmonary/non-pulmonary TB. Interventions are highlighted with vertical lines, with linetype denoting the type of intervention, more information on each intervention is available in the corresponding table. Figure produced using `tbinenglanddataclean`", out.extra = ""}
plot_historic_tb_ew() + theme(legend.position = "top")+ 
  theme(text = element_text(size = 12))
```

@@

```{r plot-prop-pul, echo = FALSE, message = FALSE, warning = FALSE, dpi = 300, fig.height = 12, fig.width = 12, fig.scap = "From 1913 until 1981 the figure shows the proportion respiratory vs. non-respiratory cases and from 1982 it shows the proportion of pulmonary vs. non-pulmonary TB.", fig.cap = "From 1913 until 1981 the figure shows the proportion respiratory vs. non-respiratory cases and from 1982 it shows the proportion of pulmonary vs. non-pulmonary TB. Figure produced using `tbinenglanddataclean`"}
plot_historic_prop_pul_tb() + theme(legend.position = "top") +
  theme(text = element_text(size = 12))
```

@@

The Bacillus Calmette–Guérin (BCG) vaccine was first given to humans in 1921 and remains the only licensed vaccine for TB.[@Medicine2013] The BCG vaccine is a live vaccine and was developed by weakening a strain of Mycobacterium bovis, which is commonly found in cows, over a period of 13 years.[@TheWorldHealthOrganization:2018va] Serious side effects are rare, although a small scar at the injection site is common. Initially, public acceptance was slow, with low take up until after the Second World War. However, controversy remains and several countries have scaled back, or redicted, their usage in recent years. This section details the action, effectiveness, duration of protection, effects and usage of the BCG vaccine.

@@

In addition, BCG vaccination policies have differed by the number of doses given, the method of application (although most countries now use the intradermal route), and the strain type used.[@Zwerling2011a] Policies have also changed over time within countries due to changes in evidence, global best practice, TB incidence rates and HIV incidence. This means that in order to understand the current impact of BCG vaccination in a population it is important to know the both the current vaccination policy but also historic vaccination policies.

@@

Since 2015, BCG vaccination has been included in the Cover Of Vaccination Evaluated Rapidly (COVER) programme, allowing coverage to be estimated in areas of England with universal vaccination (implemented due to high incidence rates based on WHO guidelines). Coverage for areas in England implementing targeted vaccination remains unknown. In London current coverage estimates are made by Local Authority and range from 5.3% to 92.1%.[@PHE2017] These estimates may not be reliable as COVER has only recently begun to include returns for BCG, meaning that data quality maybe poor. Prior to the switch to targeted neonatal vaccination coverage in those at school leaving age was thought to be approximately 75%.[@Sutherland1989]

>>>>

```{r plot-incidence, echo = FALSE, warning = FALSE, message = FALSE, fig.scap = "TB notifications in England and Wales from 1913 to 2017, stratified initially by respiratory/non-respiratory status and from 1982 by pulmonary/non-pulmonary TB.", fig.cap = "TB notifications in England and Wales from 1913 to 2017, stratified initially by respiratory/non-respiratory status and from 1982 by pulmonary/non-pulmonary TB. Interventions are highlighted with vertical lines, with linetype denoting the type of intervention, more information on each intervention is available in the corresponding table. Figure produced using tbinenglanddataclean", out.extra = ""}
plot_historic_tb_ew() + theme(legend.position = "top")+ 
  theme(text = element_text(size = 12))
```

@@

```{r plot-prop-pul, echo = FALSE, message = FALSE, warning = FALSE, fig.scap = "From 1913 until 1981 the figure shows the proportion respiratory vs. non-respiratory cases and from 1982 it shows the proportion of pulmonary vs. non-pulmonary TB.", fig.cap = "From 1913 until 1981 the figure shows the proportion respiratory vs. non-respiratory cases and from 1982 it shows the proportion of pulmonary vs. non-pulmonary TB. Figure produced using tbinenglanddataclean"}
plot_historic_prop_pul_tb() + theme(legend.position = "top") +
  theme(text = element_text(size = 12))
```

@@

The Bacillus Calmette–Guérin (BCG) vaccine was first given to humans in 1921 and remains the only licensed vaccine for TB.[@Medicine2013] The BCG vaccine is a live vaccine and was developed by weakening a strain of Mycobacterium bovis, which is commonly found in cows, over a period of 13 years.[@TheWorldHealthOrganization:2018va] Serious side effects are rare, although a small scar at the injection site is common. Initially, public acceptance was slow, with low take up until after the Second World War. However, controversy remains and several countries have scaled back, or retargeted, their usage in recent years. This section details the action, effectiveness, duration of protection, effects and usage of the BCG vaccine.

@@

In addition, BCG vaccination policies have differed by the number of doses given, the method of application (although most countries now use the intradermal route), and the strain type used.[@Zwerling2011a] Policies have also changed over time within countries due to changes in evidence, global best practice, TB incidence rates and HIV incidence. This means that in order to understand the current impact of BCG vaccination in a population it is important to know both the current vaccination policy but also historic vaccination policies.

@@

Since 2015, BCG vaccination has been included in the Cover Of Vaccination Evaluated Rapidly (COVER) programme, allowing coverage to be estimated in areas of England with universal vaccination (implemented due to high incidence rates and based on WHO guidelines). Coverage for areas in England implementing targeted vaccination remains unknown. In London current coverage estimates are made by Local Authority and range from 5.3% to 92.1%.[@PHE2017] These estimates may not be reliable as COVER has only recently begun to include returns for BCG, meaning that data quality may be poor. Prior to the switch to targeted neonatal vaccination coverage in those at school leaving age was thought to be approximately 75%.[@Sutherland1989]

### 2019-08-06 rmds/03-getTBinR.Rmd
updated getTBinR figures

```{r map-tb-incidence-eur, fig.scap = "Map of global TB incidence rates in 2017 as generated by getTBinR.", fig.cap = "Map of global TB incidence rates in 2017 as generated by getTBinR. Visualising the data with a map allows for spatial trends to be rapidly explored.", echo = TRUE, out.extra = ""}
map_tb_burden(metric = "e_inc_100k", verbose = FALSE)
```

@@

```{r plot-tb-overview, fig.width = 10, fig.height = 10, fig.cap = "Dot plot showing trends over time in TB incidence rates in Europe ordered by TB incidence rates in 2017. ", echo = TRUE}
plot_tb_burden_overview(metric = "e_inc_100k",
                        countries = "United Kingdom",
                        compare_to_region = TRUE,
                        interactive = FALSE,
                        verbose = FALSE)
```

@@

```{r plot-regional-summary, fig.scap = "TB incidence by region and globally as computed and visualised by getTBinR.", fig.cap = "TB incidence by region and globally as computed and visualised by getTBinR. Confidence intervals have been disabled in order to avoid obscuring the dominant trends.", echo = TRUE, out.extra = ""} 
plot_tb_burden_summary(conf = NULL, metric_label = "e_inc_100k",
                       verbose = FALSE)
```

@@

```{r plot-tb-burden, fig.scap = "TB incidence rates over time, with confidence intervals, in the UK. As produced by getTBinR.", fig.cap = "TB incidence rates over time, with confidence intervals, in the UK. As produced by getTBinR.", echo = TRUE, out.extra = ""}
plot_tb_burden(metric = "e_inc_100k",
               countries = "United Kingdom",
               verbose = FALSE) 
```

>>>>

```{r map-tb-incidence-eur, fig.scap = "Map of global TB incidence rates in 2017 as generated by getTBinR.", fig.cap = "Map of global TB incidence rates in 2017 as generated by getTBinR. Visualising the data with a map allows for spatial trends to be rapidly explored.", echo = TRUE, out.extra = "", fig.height = 4}
map_tb_burden(metric = "e_inc_100k", verbose = FALSE) +
  theme(text = element_text(size = 12))
```

@@

```{r plot-tb-overview, fig.cap = "Dot plot showing trends over time in TB incidence rates in Europe ordered by TB incidence rates in 2017. ", echo = TRUE}
plot_tb_burden_overview(metric = "e_inc_100k",
                        countries = "United Kingdom",
                        compare_to_region = TRUE,
                        interactive = FALSE,
                        verbose = FALSE) +
  theme(text = element_text(size = 12))
```

@@

```{r plot-regional-summary, fig.scap = "TB incidence by region and globally as computed and visualised by getTBinR.", fig.cap = "TB incidence by region and globally as computed and visualised by getTBinR. Confidence intervals have been disabled in order to avoid obscuring the dominant trends.", echo = TRUE, out.extra = ""} 
plot_tb_burden_summary(conf = NULL, metric_label = "e_inc_100k",
                       verbose = FALSE) +
  theme(text = element_text(size = 12))
```

@@

```{r plot-tb-burden, fig.scap = "TB incidence rates over time, with confidence intervals, in the UK. As produced by getTBinR.", fig.cap = "TB incidence rates over time, with confidence intervals, in the UK. As produced by getTBinR.", echo = TRUE, out.extra = ""}
plot_tb_burden(metric = "e_inc_100k",
               countries = "United Kingdom",
               verbose = FALSE) +
  theme(text = element_text(size = 12))
```
